Bayesian Sparsification of $\cplx$-valued networks

### 2019-11-05 paper/paper.tex
compressing bayesian dropout section

Kullback-Leibler divergence, defined for $q(\omega) \ll p(\omega)$ as
$$
KL(q\| p)
  = \mathbb{E}_{\omega \sim q}
    \log\frac{q(\omega)}{p(\omega)}
  \,, $$
is the natural choice for $\rho$, despite it not being a metric and requiring distributions
with nested support. The main reason is that it satisfies the following identity, which when
everything works, holds for any $q$ and $p$:
\begin{equation} \label{eq:kl-div-master}
  \log p(D)
    % = \mathbb{E}_{q(\omega)} \log \frac{p(D, \omega)}{p(\omega \mid D)}
    % = \mathbb{E}_{q(\omega)} \log \frac{
    %   p(D \mid \omega) \pi(\omega)
    % }{p(\omega \mid D)}
    % = \mathbb{E}_{q(\omega)} \log \frac{
    %   p(D \mid \omega) \pi(\omega)
    % }{q(\omega)}
    % + \mathbb{E}_{q(\omega)} \log \frac{q(\omega)}{p(\omega \mid D)}
    % = \mathbb{E}_{q(\omega)} \log{p(D \mid \omega)}
    % - \mathbb{E}_{q(\omega)} \log \frac{q(\omega)}{\pi(\omega)}
    % + \mathbb{E}_{q(\omega)} \log \frac{q(\omega)}{p(\omega \mid D)}
    = \mathbb{E}_{q(\omega)} \log{p(D \mid \omega)}
    - KL(q(\omega) \| \pi(\omega))
    + KL(q(\omega) \| p(\omega \mid D))
    \,.
\end{equation}
This key identity yield an equivalent problem for \eqref{eq:variational-progam}:
maximizing the \textit{evidence lower bound} (ELBO)
\begin{equation} \label{eq:elbo_general}
\begin{aligned}
  & \underset{\theta}{\text{maximize}}
    & \mathcal{L}(\theta; \phi, \lambda)
      & = \mathbb{E}_{\omega \sim q_{\theta}}
          \log p_{\phi}(D \mid \omega)
        - KL(q_{\theta}(\omega) \| \pi_{\lambda}(\omega))
        % \mathbb{E}_{\omega \sim q_{\theta}}
          % \log\frac{q_\theta(\omega)}{\pi_{\lambda}(\omega)}
  % \\
  % & &
  %     & = \mathbb{E}_{\omega \sim q_{\theta}}
  %         \log{p_{\phi}(D \mid \omega) \pi_{\lambda}(\omega)}
  %       + \mathbb{H}(q_{\theta})
  \,,
\end{aligned}
\end{equation}
where the prior and likelihood are allowed to depend on explicit parameters. The most
commonly used algorithm to optimize \eqref{eq:elbo_general} is to use EM-algorithm.
Other variational objectives are possible, provided $p(\omega \mid D)$ is probed only
through $\log p(D \mid \omega)$ and $\log \pi(\omega)$, \cite{ranganath_operator_2018}.

@@

For a particular family of posterior approximation, variational inference can be used
as a regularization method \cite{kingma_variational_2015} and as model sparsification
technique \cite{molchanov_variational_2017}. In \cite{kingma_variational_2015} Binary
dropout, DropConnect and Gaussian dropout, \cite{hinton_improving_2012,wan_regularization_2013},
\cite{srivastava_dropout_2014}, and \cite{wang_fast_2013} were reinterpreted as variational
inference methods. That study argues that the multiplicative noise, introduced by dropout,
induces a distribution equivalent to a fully factorized variational posterior of the form $
  q_\theta(\omega) = \prod_j q_{\theta}(\omega_j)
$, where $q_{\theta}(\omega_j)$ is $\omega_j = \mu_j \xi_j$ with $\xi_j \sim p(\xi_j)$
iid from some $p_\xi$. In Binary dropout with rate $p \in (0, 1)$ we have $
  p_\xi
    = \mathcal{B}\bigl(
        \{0, \tfrac1{1-p}\}, 1-p
      \bigr)
$, whereas in Gaussian dropout $
  \xi \sim \mathcal{N}(1, \alpha)
$ for $\alpha = \tfrac{p}{1-p}$. In the latter case, the \textit{dropout score} $\alpha$ acts
like a variance scaling factor, making it a natural variational parameter for the posterior
approximation, which is optimized via SGVB \eqref{eq:elbo}. Thus \cite{kingma_variational_2015}
propose Variational Dropout technique for Bayesian model regularization. The flexibility in
specifying dependence of $q_\theta$ on $\alpha$ allows imposing structure on dropout: global
rate for the whole model network, local per each component and even specific to individual
parameter.

Secondary contribution of \cite{kingma_variational_2015} has been the analysis of the local
reparameterization trick as a gradient variance reduction technique for \eqref{eq:elbo},
that is particularly suitable for Variational Dropout. The trick is to translate uncertainty
from the global parameter noise to local noise in intermediate outputs within the network.
Initially, this idea was used in \cite{who-dis-originally-22} for Gaussian dropout,
who employed the closure of Gaussians under affine transformations as a way to inject
regularizing noise. Later in \cite[app.~B]{kingma_variational_2015} it was rigorously shown
that for the case of fully factorized Gaussian dropout posterior approximations, the
gradient estimator $\nabla_\theta \tilde{\mathcal{L}}$ is more statistically efficient
and computationally fast, when the trick is used in addition to independent per sample
parameter draws, which reduce variance by decorrelating ELBO estimates within the minibatch.

Bayesian dropout technique with mean field Gaussian variational approximation is further
developed in \cite{molchanov_variational_2017} and \cite{kharitonov_variational_2018}. The
first paper studies variational dropout from a model sparsification perspective and improves
on the results of \cite{kingma_variational_2015} in two directions: the authors propose a
more accurate approximation to the KL-divergence term in \eqref{eq:elbo}, and introduce
additive noise parameterization for the factors of $q_\theta(\omega)$, $\omega = W$.

@@

In another variant of Bayesian dropout, called Automatic Relevance Determination, is developed
in the \cite{molchanov2018}. The key difference from Variational Dropout is the choice of
prior: the log-uniform factor $\pi(\omega_{ij})$ is replaced by a proper Gaussian prior
$\mathcal{N}(0, \tau^{-1}_{ij})$ with learnable precision $\tau_{ij} > 0$. This is knowns as
the Empirical Bayes method, \cite{citation_needed}, which fits a prior distribution while
performing variational inference. The optimal precision parameter is found in the closed form
by minimizing the KL-divergence term holding variational parameters $\theta$ fixed.

@@

As was demonstrated in \cite{kingma_variational_2015}, local reparameterization trick
provides both faster computations and better statistical efficiency of the SGVB gradient
estimator. We formulate the $\cplx$ version of the local reparameterization trick, which
generalizes \cite{who-dis-originally-22}.

>>>>

Kullback-Leibler divergence, defined for $q(\omega) \ll p(\omega)$ as
$$
KL(q\| p)
  = \mathbb{E}_{\omega \sim q}
    \log\frac{q(\omega)}{p(\omega)}
  \,, $$
is the natural choice for $\rho$, despite it not being a metric and requiring distributions
with nested support. The main reason is that it satisfies the following identity, which when
everything works, holds for any $q$ and $p$:
\begin{equation} \label{eq:kl-div-master}
  \log p(D)
    % = \mathbb{E}_{q(\omega)} \log \frac{p(D, \omega)}{p(\omega \mid D)}
    % = \mathbb{E}_{q(\omega)} \log \frac{
    %   p(D \mid \omega) \pi(\omega)
    % }{p(\omega \mid D)}
    % = \mathbb{E}_{q(\omega)} \log \frac{
    %   p(D \mid \omega) \pi(\omega)
    % }{q(\omega)}
    % + \mathbb{E}_{q(\omega)} \log \frac{q(\omega)}{p(\omega \mid D)}
    % = \mathbb{E}_{q(\omega)} \log{p(D \mid \omega)}
    % - \mathbb{E}_{q(\omega)} \log \frac{q(\omega)}{\pi(\omega)}
    % + \mathbb{E}_{q(\omega)} \log \frac{q(\omega)}{p(\omega \mid D)}
    = \mathbb{E}_{q(\omega)} \log{p(D \mid \omega)}
    - KL(q(\omega) \| \pi(\omega))
    + KL(q(\omega) \| p(\omega \mid D))
    \,.
\end{equation}
This key identity yield an equivalent problem for \eqref{eq:variational-progam}:
maximizing the \textit{evidence lower bound} (ELBO)
\begin{equation} \label{eq:elbo_general}
\begin{aligned}
  & \underset{\theta, \phi, \lambda}{\text{maximize}}
    & \mathcal{L}(\theta; \phi, \lambda)
      & = \mathbb{E}_{\omega \sim q_{\theta}}
          \log p_{\phi}(D \mid \omega)
        - KL(q_{\theta}(\omega) \| \pi_{\lambda}(\omega))
        % \mathbb{E}_{\omega \sim q_{\theta}}
          % \log\frac{q_\theta(\omega)}{\pi_{\lambda}(\omega)}
  % \\
  % & &
  %     & = \mathbb{E}_{\omega \sim q_{\theta}}
  %         \log{p_{\phi}(D \mid \omega) \pi_{\lambda}(\omega)}
  %       + \mathbb{H}(q_{\theta})
  \,,
\end{aligned}
\end{equation}
where the prior and likelihood are allowed to depend on explicit parameters. The most
commonly used algorithm to optimize \eqref{eq:elbo_general} is to use EM-algorithm.
Other variational objectives are possible, provided $p(\omega \mid D)$ is probed only
through $\log p(D \mid \omega)$ and $\log \pi(\omega)$, \cite{ranganath_operator_2018}.

@@

For a particular family of posterior approximation, variational inference can be used
as a regularization method \cite{kingma_variational_2015} and as model sparsification
technique \cite{molchanov_variational_2017}. \cite{kingma_variational_2015} propose
Variational Dropout, by studying Binary dropout, DropConnect and Gaussian dropout,
\cite{hinton_improving_2012,wan_regularization_2013}, \cite{srivastava_dropout_2014},
and \cite{wang_fast_2013} through the lens of variational inference methods. The study
argues that the multiplicative noise, introduced by dropout, induces a distribution
equivalent to a fully factorized variational posterior of the form $
  q_\theta(\omega) = \prod_j q_{\theta}(\omega_j)
$, where $q_{\theta}(\omega_j)$ is $\omega_j = \mu_j \xi_j$ with $\xi_j \sim p(\xi_j)$
iid from some $p_\xi$. Binary dropout with rate $p \in (0, 1)$ uses $
  p_\xi
    = \mathcal{B}\bigl(
      \{0, \tfrac1{1-p}\}, 1-p
    \bigr)
$, whereas in Gaussian dropout $
  \xi \sim \mathcal{N}(1, \alpha)
$ for $\alpha = \tfrac{p}{1-p}$. This suggests making $\alpha$ a variational parameter
and optimizing it in \eqref{eq:elbo_general} with an appropriate penalty term. Flexibility
in dependence of $q_\theta$ on $\alpha$ allows structured dropout with shared $\alpha$
across groups of parameters or sparsifying dropout, where the rate $\alpha$ is optimized
for each individual parameter, \cite{molchanov_variational_2017}.

Variational Dropout with mean field Gaussian variational approximation is further developed
in \cite{molchanov_variational_2017} and \cite{kharitonov_variational_2018}. The first paper
studies variational dropout from a model sparsification perspective and improves on the
results of \cite{kingma_variational_2015} in two directions: the authors propose a more
accurate approximation to the KL-divergence term in \eqref{eq:elbo}, and introduce additive
noise parameterization for the factors of $q_\theta(\omega)$, $\omega = W$.

@@

In another variant of Bayesian dropout, called Automatic Relevance Determination, is developed
in the \cite{kharitonov_variational_2018}. The key difference from Variational Dropout is the
choice of prior: the log-uniform factor $\pi(\omega_{ij})$ is replaced by a proper Gaussian
prior $\mathcal{N}(0, \tau^{-1}_{ij})$ with learnable precision $\tau_{ij} > 0$. This is knowns
as the Empirical Bayes method, \cite{citation_needed}, which fits a prior distribution while
performing variational inference. The optimal precision parameter is found in the closed form
by minimizing the KL-divergence term holding variational parameters $\theta$ fixed.

Second contribution of \cite{kingma_variational_2015} is the analysis of the \textit{local
reparameterization trick}, proposed in \cite{wang_fast_2013} to speed up Gaussian dropout,
as a gradient variance reduction technique for \eqref{eq:elbo}. This trick utilizes the
closure of Gaussians under affine transformations, thereby translating uncertainty from the
global parameter noise to local noise in intermediate outputs within the network.
%
\cite{kingma_variational_2015} show, that for the case of fully factorized Gaussian
posterior approximations, the trick decorrelates ELBO estimates within the minibatch,
making the gradient estimator $\nabla_\theta \tilde{\mathcal{L}}$ more statistically
efficient.

@@

As was demonstrated in \cite{kingma_variational_2015}, local reparameterization trick
provides both faster computations and better statistical efficiency of the SGVB gradient
estimator. We formulate the $\cplx$ version of the local reparameterization trick, which
generalizes \cite{wang_fast_2013}.
